Intentional release of native species undermines ecological stability

### 2022-09-22 rmd/supporting_info.Rmd
edit: tables

sim_result <- list(readRDS(here::here("output/result_ricker_2sp.rds")),
                   readRDS(here::here("output/result_ricker.rds"))) %>% 
  lapply(FUN = function(x) {
      n_sim <- x %>% 
        mutate(n_sim = n_warmup + n_burnin + n_timestep) %>% 
        pull(n_sim) %>% 
        unique()

@@

First, we analyzed a two-species community to gain insights into the model behavior and to guide the N-species analysis. Specifically, we examined the release effect with $20$ parameter values of intrinsic population growth rate $r_i$ ($0.5$ to $3.5$ with an equal interval) and carrying capacity $K$ ($50$ to $500$ with an equal interval), resulting in $400$ parameter combinations. In an ordinary Ricker model, the chosen range of $r_i$ can generate stable equilibrium, damped oscillations, limit cycle, and chaos (XXX). For simplicity, we assumed $r_1 = r_2$. We crossed this parameter setup with environmental stochasticity ($\sigma = \{0, 0.75\}$) and competition strength ($\alpha = \{0.1, 0.5\}$) and generated $1600$ simulation scenarios ($400 \times 2 \times 2$).

Under each simulation scenario, we ran `r n_sim` time steps of `r max(sim_result$n_rep)` with $100$ values of release level $R$ ($0$ to $500$ with an equal interval). We initialized the community with populations of each species drawn from a Poisson distribution with a mean of $50$. We let species grow with no release for the first `r unique(sim_result$n_warmup)` time steps to allow them reach equilibrium without intentional release (initialization period). After the initialization period, we released $R$ individuals of the enhanced species every time step over the next `r unique(sim_result$n_burnin)` time steps to reach new equilibrium with selective intentional release (burn-in period). We continued the simulation run with intentional release and saved the last `r unique(sim_result$n_timestep)` time steps, which were used to obtain the following summary statistics of the whole community $\sum_i^SN_j$: the coefficient of variation (CV), the temporal mean ($\mu$), and the SD ($\sigma$).

We subsequently analyzed a whole community model with $10$ species using important parameter combinations identified in the two-species model: intrinsic growth of enhanced species ($r_1 =$ `r unique(param_set$r1)`), average strength of interspecific competition ($\bar{\alpha}=$ `r unique(param_set$alpha)`), carrying capacity ($K=$ `r unique(param_set$k)`), and relative fitness of captive-bred individuals ($f_R=$ `r unique(param_set$phi)`). We confirmed that these parameters are influential in the ten-species community model via sensitivity analysis (**Table S2-S4**). Meanwhile, we fixed values of the following parameters: the number of species ($S=$ `r unique(param_set$n_species)`), maximum intrinsic growth rate of unenhanced species ($r_{max}=$ `r unique(param_set$r_max)`), and environmental stochasticity ($\sigma_{\epsilon}=$ `r unique(param_set$sd_env)`). This simulation setup resulted in `r nrow(param_set)` sets of parameter combinations.

As with the two-species model, we ran `r n_sim` time steps of `r max(sim_result$n_rep)` with $100$ values of release level $R$ ($0$ to $500$ with a constant interval) for each simulation scenario. Initial seed inidividuals of each species were drawn from a Poisson distribution with a mean of five. In this simulation, we repeated the seeding procedure every 10 time steps over the first `r unique(sim_result$n_warmup)` time steps to account for possible complex dynamics within a ten-species community (initialization period). The rest of simulation procedures is identical to the two-species model. We summarized values of simulation parameters in **Table S1**.

@@

`r paste(knit(text = table_stvy_chunk), collapse = '\n')`

@@

`r paste(knit(text = table_reg_chunk), collapse = '\n')`

@@

`r paste(knit(text = figure_sim_chunk), collapse = '\n')`

@@

`r paste(knit(text = figure_dyns_chunk), collapse = '\n')`

>>>>

list_sim <- list(readRDS(here::here("output/result_ricker_2sp.rds")),
                   readRDS(here::here("output/result_ricker.rds")))

list_param <- list_sim %>% 
  lapply(FUN = function(x) {
      n_sim <- x %>% 
        mutate(n_sim = n_warmup + n_burnin + n_timestep) %>% 
        pull(n_sim) %>% 
        unique()

@@

First, we analyzed a two-species community to gain insights into the model behavior and to guide the N-species analysis. Specifically, we examined the release effect with $20$ parameter values of intrinsic population growth rate $r_i$ ($0.5$ to $3.5$ with an equal interval) and carrying capacity $K$ ($50$ to $500$ with an equal interval), resulting in $400$ parameter combinations. In an ordinary Ricker model, the chosen range of $r_i$ can generate stable equilibrium, damped oscillations, limit cycle, and chaos (XXX). For simplicity, we assumed $r_1 = r_2$. We crossed this parameter setup with environmental stochasticity ($\sigma =$ `r unique(list_param[[1]]$sd_env)`) and competition strength ($\alpha =$ `r unique(list_param[[1]]$alpha)`) and generated `r nrow(list_param[[1]])` simulation scenarios ($400 \times 2 \times 2$).

Under each simulation scenario, we ran `r attr(list_param[[1]], "n_sim")` time steps of community dynamics for $100$ values of release level $R$ ($0$ to $500$ with an equal interval). We initialized populations of each species drawn from a Poisson distribution with a mean of $50$ and let species grow with no release for the first `r unique(list_sim[[1]]$n_warmup)` time steps. After the initialization period, we released $R$ individuals of the enhanced species every time step over the next `r unique(list_sim[[1]]$n_burnin)` time steps to reach new equilibrium with selective intentional release (burn-in period). We continued the simulation run with intentional release and saved the last `r unique(list_sim[[1]]$n_timestep)` time steps, which were used to obtain the following summary statistics of the whole community $\sum_i^SN_j$: the coefficient of variation (CV), the temporal mean ($\mu$), and the SD ($\sigma$).

We subsequently analyzed a whole community model with $10$ species using important parameter combinations identified in the two-species simulation: intrinsic growth of enhanced species ($r_1 =$ `r unique(list_param[[1]]$r1)`), average strength of interspecific competition ($\bar{\alpha}=$ `r unique(list_param[[1]]$alpha)`), carrying capacity ($K=$ `r unique(list_param[[1]]$k)`), and relative fitness of captive-bred individuals ($f_R=$ `r unique(list_param[[1]]$phi)`). We confirmed that these parameters are influential in the ten-species community model via sensitivity analysis (**Table S2-S4**). Meanwhile, we fixed values of the following parameters: the number of species ($S=$ `r unique(list_param[[1]]$n_species)`), maximum intrinsic growth rate of unenhanced species ($r_{max}=$ `r unique(list_param[[1]]$r_max)`), and environmental stochasticity ($\sigma_{\epsilon}=$ `r unique(list_param[[1]]$sd_env)`). This simulation setup resulted in `r nrow(list_param[[1]])` sets of parameter combinations.

As with the two-species model, we ran `r attr(list_param[[2]], "n_sim")` time steps of `r max(list_sim[[2]]$n_rep)` with $100$ values of release level $R$ ($0$ to $500$ with a constant interval) for each simulation scenario. Initial seed inidividuals of each species were drawn from a Poisson distribution with a mean of five. In this simulation, we repeated the seeding procedure every 10 time steps over the first `r unique(list_sim[[2]]$n_warmup)` time steps to account for possible complex dynamics within a ten-species community (initialization period). The rest of simulation procedures is identical to the two-species model. We summarized values of simulation parameters in **Table S1**.

@@

`r paste(knitr::knit(text = table_stvy_chunk), collapse = '\n')`

@@

`r paste(knitr::knit(text = table_reg_chunk), collapse = '\n')`

@@

`r paste(knitr::knit(text = figure_sim_chunk), collapse = '\n')`

@@

`r paste(knitr::knit(text = figure_dyns_chunk), collapse = '\n')`
